### 2017-05-15 Manuscript/ms.tex
Reframing the introduction

Dengue fever is a massive global public health burden, with millions of cases per year in Brazil alone (CITATION).  
Since the dengue virus (DENV) is vectored by the mosquito \textit{Aedes aegypti}, dengue fever is prevented primarily through mosquito control programs \cite{Achee2015}.
Even if a vaccine is developed, mosquito control is likely to remain an important component of dengue management \cite{Achee2015}.
Studies have shown that targeted mosquito control efforts can be effective in reducing disease burden.
For example contact tracing and residential spraying reduced disease risk in Cairns, Australia \cite{Vazquez-Prokopec2017}.
Though this approach was successful, it requires detailed data following individual cases, potentially limiting its utility for large-scale deployment.

Other efforts have sought to guide the deployment of control efforts through the use of fine-scale mosquito surveillance.
For instance, the "MI-Dengue" system deployed throughout several cities in Brazil uses a grid of mosquito sticky-traps (MosquiTRAPs) for spatial prioritization of source reduction and larvicide application \cite{Eiras2009}.
Comparing cities with and without this system, \cite{Pepin2013} estimated that it prevented X cases and saved X dollars in economic losses from 20XX - 20XX.
Though this system has been effective, the mosquito indices and thresholds used to prioritize control efforts remain somewhat ad-hoc.
Using such surveillance data to efficiently target limited control resources requires a detailed understanding of the relationship between the components of mosquito abundances measured by the trap and disease risk.

However, even though mosquitoes are necessary for dengue transmission, establishing a clear link between surveillance-based indices of mosquito abundance and disease risk has proven difficult thus far (CITATIONS).
Many of the attempts to establish this link have relied on tests of the ability of mosquito abundance measures to predict human disease risk in regression models.
For example, \cite{Pepin2015} found that models that included MosquiTRAP surveillance data failed to predict weekly dengue cases any better than autoregressive models that include lagged case data alone.
Evaluating the influence of mosquito abundance in this way is analogous to a test of Granger causality \cite{Granger1969}, which states that some variable $X$ (e.g. mosquito abundance) \emph{causes} $Y$ (e.g. human dengue cases) if $X$ contains unique information about $Y$ (i.e. if a model with $X$ and $Y$ provides better predictions of $Y$ than a model with $Y$ alone).
By this definition, the results of \cite{Pepin2015} suggest that mosquito abundance does not cause dengue cases (in the Granger sense).

However, as noted by Sugihara \emph{et al.} \cite{Sugihara2012a}, Granger causality assumes that $X$ and $Y$ are separable, which is not the case if $X$ and $Y$ are linked in a non-linear dynamical system.
In this case, a model with $X$ and $Y$ will not outperform the model with only $Y$, as the information contained in $X$ is already embedded in $Y$, a consequence of Taken's theorem (CITATION). 
Since dengue transmission is governed by such a non-linear dynamical system, information about the driving component of mosquito abundance should be already embedded in the time series of human case reports.
As a result, predictive utility is not an appropriate way to test the causal influence of mosquito abundance on dengue risk.
Instead, Sugihara \emph{et al.} argue that if $X$ is a driver $Y$, then you should be able to reconstruct $X$ from $Y$.

In this paper, we apply this perspective to the problem of linking mosquito abundance to disease risk.
To test whether or not mosquito abundance, as measured through surveillance efforts, drives dengue dynamics, we ask whether we can reconstruct a time series of mosquito surveillance data from a time series of case reports.
We approach this as an inverse problem: we first posit a mechanistic model of dengue transmission and then develop a Bayesian statistical framework to estimate the mosquito abundance time series that is consistent with that model and the time series of case reports.
We then evaluate the causal link between mosquito abundance and case reports by using posterior predictive checks to compare the estimated and observed mosquito time series.

>>>>

Dengue fever is a massive global public health burden, with millions of cases per year in Brazil alone (CITATION).  
Since the dengue virus (DENV) is vectored by the mosquito \textit{Aedes aegypti}, dengue fever is prevented primarily through mosquito surveillance and control programs \cite{Achee2015}.
Deploying these mosquito control measures effectively and efficiently requires a detailed understanding of how mosquito surveillance indices map to human disease risk.
For example contact tracing and residential spraying reduced disease risk in Cairns, Australia \cite{Vazquez-Prokopec2017}.
This approach ensures that the mosquitoes targeted by control are epidemiologically relevant, but the detailed data it requires potentially limits its utility for large-scale deployment.

More commonly, cities deploy mosquito control based on trap-based mosquito surveillance data.
For instance, the "MI-Dengue" system deployed throughout several cities in Brazil uses a grid of mosquito sticky-traps (MosquiTRAPs) for spatial prioritization of source reduction and larvicide application \cite{Eiras2009}.
Comparing cities with and without this system, \cite{Pepin2013} estimated that it prevented X cases and saved X dollars in economic losses from 20XX - 20XX.
Though this system has been effective, the mosquito indices and thresholds used to prioritize control efforts remain somewhat ad-hoc.
Using such surveillance data to efficiently target limited control resources requires a more detailed understanding of the relationship between the components of mosquito abundance measured by traps and disease risk.

Many of the attempts to establish this relationship statistically have relied on tests of the ability of mosquito abundance measures to predict human disease risk in regression models.
For example, \cite{Pepin2015} found that models that included MosquiTRAP surveillance data failed to predict weekly dengue cases any better than models that include lagged case data alone.
However, these models fail to account for the mechanistic, non-linear relationship between mosquito population dynamics and dengue epidemiology.
In particular, the cycle of transmission between humans and dengue is controlled not just by mosquito population size, but also by the emergence of new susceptible adults and the mortality rate.

In this paper, we seek to connect mosquito surveillance data with human case reports through the use of a Bayesian, mechanistic model.
We first propose a differential equation model for dengue epidemiology.
We then embed this model into a Bayesian statistical framework that allows us to estimate latent time series of mosquito demographic parameters from time series of mosquito trap counts and reported cases of dengue fever.
Lastly, we explore the ability of this fitted model to predict annual dengue outbreaks and its utility for targeting control efforts.
